Adopting agile software development in the large organizations: A systematic literature review

### 2014-06-06 journal/article.tex
Research method: editing, additions

@@

A key question in the study was the definition of \emph{large scale}. Large
scale requires additional coordination and makes communication difficult
compared to software development in small teams.

@@

The research was conducted as an application of Kitchenham's
\cite{Kitchenham2007} guidelines for systematic literature review. The selection
of primary studies was done as a keyword based database search and manual
filtering of included studies. The filtering process was executed independently
by two researchers (Dikert, Paasivaara). The selected primary studies were coded
for data extraction and analysis. The entire process was audited and mentored by
a third researcher (Lassenius).

@@

Before proceeding with the database search a number of preliminary searches were
performed. The purpose of the preliminary searches was to create a benchmark of
potentially relevant papers that should be matched by the actual database
search, and to evaluate different search strings. We started by examining top
ranked hits by trivial keywords that the more complex final search string might
miss. Initial searches were made using keywords that were as general as
possible, including ``agile transformation'' and ``large scale agile''.
Secondly, a trial keyword search was done in the selected databases (listed in
table \ref{table:databases}) using the following search string:

@@

The selection of studies was primarily based on a database search. The initial
step in the database search was constructing a search string which yielded a
reasonably large result set. The search result was then refined in two steps, by
first filtering out abstracts and finally full text filtering. The study
selection process starting from the database search is outlined in figure
\ref{fig:selectionprocess}. The result of the database search was still amended
with some results from the preliminary search and relevant references from all
examined papers.

\begin{figure}
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/research_process.pdf}
    \caption{Database search as study selection process}
    \label{fig:selectionprocess}
  \end{center}
\end{figure}

A keyword search in online databases was chosen as the primary means to find
potential primary studies. The databases included in the search are listed in
Table \ref{table:databases}. All databases supported use of complex Boolean
logic in searches, which allowed us flexibility in constructing the search
strings.

@@

The database keyword search matched 1874 unique papers. The abstracts of these
papers were categorized independently by both Dikert and Paasivaara into three
categories: include, exclude, and uncertain. The result was combined by marking
abstracts both researchers agreed on for inclusion or exclusion respectively.
The inclusion of disagreed or uncertain abstracts was resolved through
discussion. At this stage papers were excluded only if both researchers deemed
it clearly irrelevant, including any uncertain cases for full text filtering. As
a result 170 papers were selected for full text filtering.

We also evaluated the keyword search against the benchmark created in the
preliminary search step, concluding that 75 of the 117 preliminarily selected
papers matched. The missed 32 preliminary papers were examined, resulting in
including 3 additional papers as primary sources.

@@

After the full text filtering step the references of all papers were also
examined for relevance. Most case study papers used references very scarcely,
mostly referencing well known descriptions of agile methods. We included 2
referenced papers in full text analysis.

Finally, counting 47 papers from the full text filtering step, 3 papers from
preliminary searches, and 2 papers from references, 52 papers were selected
as primary studies. The full selection process is illustrated in figure
\ref{fig:selectionprocess_full}.

In several cases there were more than one primary study presenting the
transformation of the same organization. Even if the transformation of a single
organization was described in many studies all sources that passed the inclusion
criteria were included. Studies presenting the same organization were treated
as one unit so that we could gain as much insight from each organization as
possible. There were also a few papers that presented multiple case studies, and
in those cases we treated each studied organization individually. As a result 42
unique organizations were identified in the primary studies. We use the term
\emph{study} to refer to the primary study publications, and the term
\emph{case} to refer to an individual case organization that may be described in
several different studies.

% -----------------------

% - pitää olla systemaattinen
% - pitää olla toistettava
% ==> Perustuu enimmäkseen siihen että on 2 tutkijaa toistamassa
% 
% Jos koko on epäselvä:
% --> Tutkijat päättävät keskustelun ja kvalitatiivisen harkinnan perusteella
% --> Tähän tarvitaan esimerkit, milloin on otettu mukaan, milloin on karsittu
%     pois

@@

The primary research for this literature review consists almost exclusively of
industry experience reports. This may be seen as a factor lessening the
evidence.
--> Why did we do this anyway?
--> Case studies with research method studying organizational transformations in
    software industry are very scarce.
--> Large organizational changes are hard to measure experimentally (controlled
    experiments?). Also due to this any student research was excluded.

The reason for deviating study assessment
--> Original research consisted almost solely of experience reports
--> The fact that experience reports have a subjective standpoint makes it
    unreliable to make quantitative conclusions based on the data
--> We also decided to include separate studies describing the same organization.
    This was useful as it deepened the understanding of each case, and several
    large organizations have produced more than one description of their agile
    transformation.

The fact that the study is based on a quantitative basis
--> Experience reports have a subjective bias
--> The data synthesis must not rely on quantitative measures. This means that
    we also considered topics raised by only a marginal number of reports.
--> It may still be interesting to make some quantitative observations such as
    a particular problem being reported in the majority of reports.

It was quite common that a single transformation case was described by many
papers. For example, one paper would highlight the viewpoint of developers
\cite{Fry2007}, and another would consider the transformation from user
experience designers' point of view \cite{Federoff2009}. Instead of using the
most complete paper as suggested by Kitchenham \cite{Kitchenham2007}, we
combined the results presented in each paper and considered the case as a single
unit. By doing this we avoided bias caused by duplicate publications, and were
enabled to have a more detailed view of the case organization.

\subsubsection{Coding of primary documents}

@@

@@

>>>>

The research process is presented starting from the research questions. After
that we present the definition of large scale which this research builds upon.
The bulk of this section concentrates on presenting the steps of the research
process, starting from the definition of inclusion criteria, continuing with
describing how the primary studies were selected, and finally presenting the
method for synthesizing the primary studies. The section concludes with a
discussion on limitations of the method.

@@

A key question in this study was the definition of \emph{large scale}. Large
scale requires additional coordination and makes communication difficult
compared to software development in small teams. One of the motivations for this
research was to identify phenomena that appear particularly in large-scale
transformations, and which may not be present in small organizations.

In this section we first present the definition of large scale.

@@

The research was conducted as an application of Kitchenham's
\cite{Kitchenham2007} guidelines for systematic literature review. Selection of
primary studies was done as a keyword based database search and manual filtering
of included studies. The filtering process was executed independently by two
researchers (Dikert, Paasivaara). The selected primary studies were coded for
qualitative data extraction and analysis. The entire process was audited and
mentored by a third researcher (Lassenius).

@@

Before proceeding with identifying the primary studies a number of preliminary
searches were performed. The purpose of the preliminary searches was to create a
benchmark of potentially relevant papers that should be matched by the actual
search, and to evaluate different search strings. We started by examining top
ranked hits by trivial keywords that the more complex final search string might
miss. Initial searches were made using keywords that were as general as
possible, including ``agile transformation'' and ``large scale agile''.
Secondly, a trial keyword search was done in the selected databases (listed in
table \ref{table:databases}) using the following search string:

@@

\subsubsection{Identification of primary studies}

The gathering of potential primary studies was based on a search in online
databases. The databases searched are listed in Table \ref{table:databases}. All
databases supported use of complex Boolean logic in searches, which allowed us
flexibility in constructing the search strings. The main task in the database
search was to construct a search string which yielded a reasonably large result
set. The set of matched primary studies was later ammended with a small number
of studies found in references of the matched papers.

@@

The database keyword search matched 1875 unique papers. The results of the
search were gathered for the selection step using reference management software.

The set of potential primary studies identified by the database search was
refined in two steps, by first filtering out abstracts and finally full text
filtering. The study selection process starting from the database search is
outlined in figure \ref{fig:selectionprocess_full}. The result of the database
search was amended with some results from the preliminary search and relevant
references from all examined papers.

The database keyword search had matched 1875 unique papers. The abstracts of
these papers were categorized independently by both Dikert and Paasivaara into
three categories: include, exclude, and uncertain. There were 1578 exclusions
and 62 inclusions that both researchers agreed on. The inclusion decisions
for the 235 abstracts with disagreement or uncertainty were resolved through
discussion. At this stage papers were excluded only if both researchers deemed
it clearly irrelevant, including any uncertain cases for full text filtering. As
a result 170 papers were selected for full text filtering.

@@

We evaluated the result of the database keyword search against the benchmark
created in the preliminary search step. We concluded that 75 of the 117
preliminarily selected papers were matched by the database keyword search.
The missed 32 preliminary papers were examined, resulting in including 3
additional papers as primary sources.

In parallell with the full text filtering step the references of all papers were
also examined for relevance. Most papers used references very scarcely, typicaly
referencing well known descriptions of agile methods. We included 2 referenced
papers in full text analysis.

Finally, counting 47 papers from the full text filtering step, 3 papers from
preliminary searches, and 2 papers from references, 52 papers were selected
as primary studies.

\subsubsection{Handling of duplicate reports on a single case}

In several cases there were more than one primary study presenting the
transformation of the same organization. Duplicate descriptions of the same
organization focused typically on different aspects. For example, one paper
would highlight the viewpoint of developers \cite{Fry2007}, and another would
consider the transformation from user experience designers' point of view
\cite{Federoff2009}.

Even if the transformation of a single organization was described in many
studies, all sources that passed the inclusion criteria were included. Studies
presenting the same organization were treated as one unit so that we could gain
as much insight on each organization as possible. Conversely, there were also a
few papers that presented multiple case studies, and in those cases we treated
each studied organization individually.

As a result 42 unique organizations were identified in the primary studies. We
use the term \emph{study} to refer to the primary study publications, and the
term \emph{case} to refer to an individual case organization that may be
described in several different studies.

@@

The primary research for this literature review consists almost exclusively of
industry experience reports. There were only 6 case studies with a research
method, and observations on transformation were presented only as a minor part
in these studies. Based on this finding we concluded that case studies
presenting viewpoints on organizational transformations are very scarce in
software industry. We deemed that the results would be distorted heavily and
many valuable studies would be left out if a strict quality assessment would be
part of the inclusion criteria. As a result we decided to include all experience
reports, regardless of the perceived objectivity.

The absence of research methods and subjective viewpoints of experience reports
can be seen as a factor lessening the evidence. Because of this we were
refrained from making quantitative interpretations of the data, but based the
results around describing qualitative observations. Regardless of the bias of
the primary studies, we think that the concepts that emerged from the source
material should not be questioned on a fundamental level. The overall existence
of organizational phenomena should not be questioned based on possible author
bias in experience reports.

% --> Large organizational changes are hard to measure experimentally (controlled
%     experiments?). Also due to this any student research was excluded.

% --> It may still be interesting to make some quantitative observations such as
%     a particular problem being reported in the majority of reports.

% * The primary studies did not make distinction between method, results, and
%   discussion
% 
% * The studies neither did any implications for generalizability, but rather
%   presented narratives or experiences. Some cases presented lists of
% %   recommendations.

As presented in the previous section, we allowed several descriptions of the
same case to be included. Instead of using the most complete paper as suggested
by Kitchenham \cite{Kitchenham2007}, we combined the results presented in each
paper and considered the case as a single unit. Keeping in mind the potential
bias caused by duplicate publications, including all papers enabled us to have a
more detailed view of the case organization.

\subsubsection{Coding of primary documents}
\label{sec:coding}

@@

\subsubsection{Synthesis of primary studies based on coding}

The synthesis of findings was created based on the result of coding presented in
section \ref{sec:coding}. An initial organization of codes into categories was
made based on the labels of the codes. Each codes was linked to a single
research question, so the categorization of codes was made separately within
the topic of each research question. The categories of codes were labeled
appropriately to give a high level description of the linked codes. Some codes
having minor importance and few quotations were left out.

After assigning each code to a unique category, the content of the categories
was studied. Each category was studied by reading through each quotation of each
code included. Typically the quotations were displayed and examined in their
original context, considering also surrounding paragraphs of the quotation.
Notes were made on each quotation presenting noteworthy observations. Based on
the notes, an accurate description emerged for each code.

As an accurate description had been created for each code, the initially defined
categories could be refined. Some codes were recategorized, and the definitions
of a few categories were revisited, until a final ordering was reached. The
results are presented according to the final categorization.

% * If a code had more quotations there were more observations made (longer
%   result text)

@@

Bias  --  systematic error

Internal quality  --  validity

External quality  --  generalizability
